 

### 2022-11-08 code/Matrinem_data_analysis.Rmd

# MATRINEM - Attempts at creating a NEC mouse model

## The research questions

Experiment no. 4 was the goal, but before we could test our overall hypothesis we needed to find out which combination of antibiotics and feeding scheme would results in the most NEC-like phenotype.

These research questions were investigated in 3 pre-experiments (PreExp 1-3), each with their own research questions.

### **Overview of all experiments**

+-----------+-------------------------------------------------------------------------------+----------+----------+---------+
|           | PreExp 1                                                                      | PreExp 2 | PreExp 3 | NEC-FvT |
+===========+===============================================================================+==========+==========+=========+
| **Aim**   | To test feasibility of formula feeding with PICC line                         |          |          |         |
+-----------+-------------------------------------------------------------------------------+----------+----------+---------+
| Protocols | Testing a broad spectrum antibiotics mix in combination with formula feeding. |          |          |         |
|           |                                                                               |          |          |         |
|           | 4 groups                                                                      |          |          |         |
|           |                                                                               |          |          |         |
|           | 1.  CON-BF                                                                    |          |          |         |
|           | 2.  CON-FF                                                                    |          |          |         |
|           | 3.  AB-BF                                                                     |          |          |         |
|           | 4.  AB-FF                                                                     |          |          |         |
+-----------+-------------------------------------------------------------------------------+----------+----------+---------+
| N         | 38 (32 survived)                                                              |          |          |         |
+-----------+-------------------------------------------------------------------------------+----------+----------+---------+

@@

#Load in the data

```{r include=FALSE}
preexp_1 <- read_excel("data/processed/weight_development.xlsx") %>% 
  filter(exp_number == "1") 
```

@@

#### **Rationale:**

*Chen et al.*, showed that Maternal Antibiotics Treatment (**MAT**) resulted in mild NEC-like symptoms of mice offspring. We believed that MAT combined with formula feeding (**FF**) of offspring would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

#### **Aim:**

To test if it was possible to keep mice separated from the mother alive from postnatal day 3 and 48 hours onwards.

#### **Research Questions:**

##### Can we keep offspring alive with formula feeding?

@@

```{r}
#___________ Check cytokine distribution for normality, filter significant ones _____________________
normality_check <- cytokines_1_long %>% 
  group_by(cytokines) %>% 
  shapiro_test(cytokine_concentration) %>% 
  select(cytokines, p)
#___________________________________________________________

>>>>

@@

#Load in the data
```{r exp1-weight, include=FALSE}
#____Load overview of sample IDs______ 
pups <- read_excel("data/processed/weight_development.xlsx")

#____ Separate into experimental rounds ____#
preexp_1 <- pups %>% 
  filter(exp_number == "1")

exp1_survived <- preexp_1 %>% 
  filter(!is.na(bodyweight_baseline_plus_48h)) %>% 
  nrow()
#____ ____#
```

```{r exp1-groups}
#____Separate into groups ______ 
AB_FF <- nrow(preexp_1 %>% 
  filter(group2 == "AB-FF"))

AB_BF <- nrow(preexp_1 %>% 
  filter(group2 == "AB-BF"))

CON_FF <- nrow(preexp_1 %>% 
  filter(group2 == "CON-FF"))

CON_BF <- nrow(preexp_1 %>% 
  filter(group2 == "CON-BF")) 
#____ ____#

Survival rate: `r exp1_survived` of `r nrow(preexp_1)`(`r (exp1_survived / nrow(preexp_1))*100`%)

@@

@@

```{r}
#___________ Check cytokine distribution for normality, filter significant ones _____________________
normality_check <- cytokines_1_long %>% 
  group_by(cytokines) %>% 
  shapiro_test(cytokine_concentration) 
#___________________________________________________________

### 2022-11-08 code/Round 1/Cytokines adjusted.Rmd

First you load the data and check the structure. Make sure cytokine concentrations are all numeric (<dbl>). If not, mutate non-numeric columns with mutate_if(is_character).  
```{r load data, echo=FALSE}
# Import dataset
cytokines <- read_excel("../../data/processed/cytokines.xlsx")
glimpse(cytokines)

cytokines_adjusted <- cytokines %>% 
  pivot_longer(.,
               cols = INFg:TNFa ,
               names_to = "variables",
               values_to = "values"
  ) %>% 
  mutate(
    group = recode_factor(  #Change the order of the groups to desired plot order
      group,
      "CON-BF"= "CON-BF",
      "CON-FF" = "CON-FF",
      "AB-BF"= "AB-BF",
      "AB-FF"= "AB-FF"
      )) %>% 
  mutate(values = if_else(tissue == "ileum", values/tissue_weight_mg, values)) %>%
  pivot_wider(
    id_cols = c(
      sample_id,exp_number, group, received_antibiotics,maternal_treatment, tissue,tissue_weight_mg),
    names_from = variables, 
    values_from = values)

#____Pivot longer____
cytokines_long <- cytokines_adjusted %>% 
  pivot_longer(.,
               cols = INFg:TNFa ,
               names_to = "variables",
               values_to = "values"
  )
#pivot_wider(id_cols = c(sample_id,exp_number, group, received_antibiotics,maternal_treatment, #tissue,tissue_weight_mg, adjusted_values), names_from = variables, values_from = values)
```

@@

#Density plots. Works
histogram <- cytokines_long %>% 
  filter(exp_number == 1) %>% 
  filter(tissue == "ileum") %>%
  filter(variables != "INFg") %>% 
  drop_na(adjusted_values) %>% 
  ggplot(aes(x = adjusted_values, fill = group)) + 
  geom_density() +
  facet_wrap(~ variables,
             scales = "free")

@@

#Jitter plots. Works
jitter_plot <- cytokines_long %>% 
  filter(exp_number == 1) %>% 
  filter(tissue == "ileum") %>%
  #filter(variables != "INFg") %>% 
  drop_na(adjusted_values) %>% 
  ggplot(aes(x = group, y = adjusted_values, color = group)) +
  geom_jitter() +
  facet_wrap(~ variables,
             scales = "free")

log10_jitter_plot <- cytokines_long %>% 
  filter(exp_number == 1) %>% 
  filter(tissue == "ileum") %>%
  filter(variables != "INFg") %>% 
  drop_na(adjusted_values) %>% 
  mutate(adjusted_values = log10(adjusted_values)) %>% 
  ggplot(aes(x = group, y = adjusted_values, color = group)) +
  geom_jitter() +
  facet_wrap(~ variables,
             scales = "free") 
```

@@

tbl1 <- cytokines %>% 
  filter(exp_number == "1") %>% 
  filter(tissue == "ileum") %>% 
  select(group,INFg:TNFa) %>% 
  tbl_summary(by = group,
              digits = everything() ~ c(2,1,1),
              missing = "no",
              include = !"INFg",
             # statistic = list(all_continuous() ~ "{mean} ({sd})")  #Un-comment this line to get mean+sd. 
             #%>%
statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
  add_p() %>% 
  #modify_footnote("Exp. 1") %>% 
  modify_header(
    label ~ "Cytokines") %>% 
  modify_spanning_header((c("stat_1", "stat_2", "stat_3", "stat_4") ~ "Treatment groups (N total = {N})"))

>>>>

First you load the data and check the structure. Make sure cytokine concentrations are all numeric (<dbl>). If not, mutate non-numeric columns with mutate_if(is_character).  
```{r load data, echo=FALSE}
# Import dataset
cytokines <- read_excel("data/processed/cytokines.xlsx")

tissue_weight_destribution <- cytokines %>%   # run normality test on tissue weight
  filter(tissue == "ileum") %>% 
  filter(exp_number == "1") %>% 
  group_by(group) %>% 
  shapiro_test(tissue_weight_mg)

# group  variable         statistic      p
#   <chr>  <chr>                <dbl>  <dbl>
# 1 AB-BF  tissue_weight_mg     0.917 0.404 
# 2 AB-FF  tissue_weight_mg     0.953 0.686 
# 3 CON-BF tissue_weight_mg     0.771 0.0138
# 4 CON-FF tissue_weight_mg     0.812 0.144

#look at tissue weight distribution
cytokines %>% 
  filter(tissue == "ileum") %>% 
  filter(exp_number == "1") %>% 
  ggplot(aes(x = group, y = tissue_weight_mg, fill = group)) + 
  geom_boxplot() +
  geom_point()
```
<!-- Looks fine. Go ahead and  adjust-->

```{r adjust for tissue weight , echo=FALSE}
cytokines_adjusted <- cytokines %>% 
  pivot_longer(.,
               cols = INFg:TNFa ,
               names_to = "variables",
               values_to = "values"
  ) %>% 
  mutate(
    group = recode_factor(  #Change the order of the groups to desired plot order
      group,
      "CON-BF"= "CON-BF",
      "CON-FF" = "CON-FF",
      "AB-BF"= "AB-BF",
      "AB-FF"= "AB-FF"
      )) %>% 
  mutate(values = if_else(tissue == "ileum", values/tissue_weight_mg, values)) %>% #Adjust for tissue weight
  pivot_wider(
    id_cols = c(
      sample_id,exp_number, group, received_antibiotics,maternal_treatment, tissue,tissue_weight_mg),
    names_from = variables, 
    values_from = values)

#____Pivot longer____
cytokines_long <- cytokines_adjusted %>% 
  pivot_longer(.,
               cols = INFg:TNFa ,
               names_to = "variables",
               values_to = "values"
  )
```

@@

#Density plots. Works
histogram <- cytokines_long %>% 
  filter(exp_number == 1) %>% 
  filter(tissue == "ileum") %>%
  filter(variables != "INFg") %>% 
  drop_na(values) %>% 
  ggplot(aes(x = values, fill = group)) + 
  geom_density() +
  facet_wrap(~ variables,
             scales = "free")

@@

#Jitter plots. Works
jitter_plot <- cytokines_long %>% 
  filter(exp_number == 1) %>% 
  filter(tissue == "ileum") %>%
  #filter(variables != "INFg") %>% 
  drop_na(values) %>% 
  ggplot(aes(x = group, y = values, color = group)) +
  geom_jitter() +
  facet_wrap(~ variables,
             scales = "free")

log10_jitter_plot <- cytokines_long %>% 
  filter(exp_number == 1) %>% 
  filter(tissue == "ileum") %>%
  filter(variables != "INFg") %>% 
  drop_na(values) %>% 
  mutate(adjusted_values = log10(values)) %>% 
  ggplot(aes(x = group, y = values, color = group)) +
  geom_jitter() +
  facet_wrap(~ variables,
             scales = "free") 
```

@@

tbl1 <- cytokines %>% 
  filter(exp_number == "1") %>% 
  filter(tissue == sample_tissue) %>% 
  select(group,INFg:TNFa) %>% 
  tbl_summary(by = group,
              digits = everything() ~ c(2,1,1),
              missing = "no",
              include = !"INFg",
             # statistic = list(all_continuous() ~ "{mean} ({sd})")  #Un-comment this line to get mean+sd. 
             #%>%
statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
  add_p() %>% 
  #modify_footnote("Exp. 1") %>% 
  modify_header(
    label ~ "Cytokines") %>% 
  modify_spanning_header((c("stat_1", "stat_2", "stat_3", "stat_4") ~ "Treatment groups (N total = {N})"))

### 2022-11-08 submission/thesis.Rmd

```{r uni_logo, echo=F, out.width="95%"}
knitr::include_graphics("img/sund.pdf")
```

\Large
{\bf The University Of Edinburgh}

\Large
{\bf School Of Geosciences}

@@

\doublespacing
{\bf COMPARISON OF TOP-DOWN AND BOTTOM-UP APPROACHES ON SPECIFIC LEAF AREA PATTERNS, \\AT GLOBAL, LATITUDINAL, AND BIOME SCALES}

@@

{\bf ANNA CHIRUMBOLO}

@@

in partial fulfilment of the requirement \\for the degree of BSc with Honours \\in Ecological and Environmental Sciences

@@

@@

```{r results, child= 'results.Rmd'}
```

>>>>

\begin{centering}
```{r uni-logo-top, echo=F, out.width="80%"}
knitr::include_graphics("img/sund.png")
```

\Large
{\bf The University Of Copenhagen}

\Large
{\bf Factuly of Health}

@@

\doublespacing
{\bf MATERNAL ANTIBIOTICS TREATMENT AND FORMULA FEEDING \\ATTEMPTS AT CREATING A NEC MOUSE MODEL}

@@

{\bf KAARE D. TRANÆS}

@@

Master thesis project \\MSc Human Biology \\SUND

@@

```{r uni-logo-bottom, echo=F, out.width="60%"}
knitr::include_graphics("img/absalon_watermark.png")
```

@@

#```{r results, child= 'results.Rmd'}
#```
